Testing Quasar Unification: I. Radiative Transfer in Clumpy Winds

### 2015-12-15 paper.tex
about to shorten angular distribution discussion

\title
{
% Modelling quasar outflows:
% clumpy winds and broad emission lines
Testing Quasar Unification with Clumpy Wind Models
%Unifying quasars with clumpy wind models
}

@@

Overall, our work suggests that geometric unification

@@

pose a number of difficult challenges to models in which an equatorial outflow

@@

Our previous modelling efforts, we have assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
Implementing a treatment of clumping is challenging, for
two main reasons. First, the physical scale lengths and density contrasts 
associated with these parameters are not well-constrained from observations.  
Second, there are significant computational difficulties associated with adequately
resolving and realistically modelling a series of small scale, high density
regions with a MCRT code. 
% Second, the addition of multiple additional degrees of
% freedom in the model results in significantly wider parameter space.
% Unfortunately, the physical scale lengths and density contrasts 
% associated with these parameters are not well-constrained from observations.  
% {\bf KSL: Reverse the order of these difficulties.  The most important is that it is not clear what kind of clumping an AGN wind has; the computational problem is a secondary one JM: Done..}

Given the lack of knowledge about the actual type of clumping, we have adopt a simple approximation
used successfully in stellar wind modelling, known as 
{\em microclumping} \citep{hamann1998,hamann2008}(MORE REFS), 
which minimizes the computational difficulties. 
The underlying assumption in microclumping is that clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `volume filling factor', $f_V$. 
The intra-clump medium is assumed to be a vacuum, so the 
density of the clumps is then multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as 
\begin{equation}
\kappa = f_V \kappa_C(D);~~\epsilon = f_V \epsilon_C(D).
\end{equation}
Here the subscript $C$ denotes that the quantity is calculated using the 
enhanced density in the clump. The resultant effect is that, {\em for fixed temperature},
 processes that are {\em linear} in density, such as electron scattering, are unchanged, 
as $f_V$ and $D$ will cancel out. However, any quantity that scales with the {\em square} of density, 
such as collisional excitation or recombination, will increase by a factor of $D$.
However, in the event that one of these

Clumping the wind has an important effect on the ionization state and has
been proposed as a solution to the so-called `over-ionization problem' in 
disc winds \citep{hamann2013}. This is the main motivation for incorporating microclumping
into our model. This treatment is necessarily simple; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless,  this parameterization allows simple estimates
of the effect clumping might have on the ionization state and emergent 
line emission. 
% It is also encouraging that microclumping has been used 
% successfully in fits to O-star wind spectra \citep{hillier1991eswingsmodel}.

@@

We then evaluated these models based on the following criteria:
\begin{itemize}
\item Does the model avoid over-ionization and thus produce UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles (e.g. Knigge et al. 2010)?
\item Do the synthetic spectra show  line emission emerge at low inclinations, with $EW\sim40\AA$ in \civ \citep[e.g. ][]{shen2011}?
\item Do H recombination lines appear in the spectra, $EW\sim50\AA$ in \la 
\citep[e.g. ][]{dipompeo2012}?
\item Do the spectra produce \mg\ and \al\ (LoBAL) absorption 
features with $BI > 0$ at a small subset of BAL angles?
\item Does the model spectra resemble quasar composite spectra?
\end{itemize}
Where $BI$ is the `Balnicity Index' (Weymann et al. 1991), given by
\begin{equation}
BI = \int^{25000}_{3000} \left( 1 - \frac{f(v)}{0.9} \right) dv.
\end{equation}
The constant $C=0$ everywhere, unless the normalized flux
has satisfied $f(v)<0.9$ continuously for at least $2000$~km~s$^{−1}$, 
whereby $C$ is set to $1$.

@@

\section{Results From A Fiducial Model}

Here we describe the results from an fiducial model,
and discuss these results in the context of the criteria 
presented in section 3.4. The parameters of this model are shown in table~2.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk. In this section, we examine the physical 
conditions of the flow, and present the synthetic spectra, before comparing
the X-ray properties of this particular model to samples of
quasars and luminous AGN. 
% We also examine trends with inclination in the synthetic spectra, 
% both in terms of the range of ionization states of the absorption lines and equivalent widths of the emission lines.

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Next-generation Model Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{erg~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.6^*$ \\
$f_V$ 	&	 $0.01^*$  \\
$n_x$ 	&	 $100$  \\
$n_z$ 	&	 $200$  \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
% $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters 
used in the model, as defined in the text and figure 1.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk.}
\label{wind_param}
\end{table}

@@

Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clump densities 
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane, 
where $n_H$ is the local number density of H.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity equal to the escape velocity at the streamline base
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration results in
a wind that exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
This is illustrated in figure~\ref{fig:wind} 
by the panels showing the ion fraction $F=n_j/n_{tot}$ of some important ions.
With a clumped wind, we are able to produce the range of ionization states observed
in quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{erg~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

The total line luminosity also increases dramatically compared to H13. 
This is due to heating of the plasma by the increased the increased X-ray 
luminosity heating the plasma, while clumping prevents over-ionization.
This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the incident X-rays.
The radii of these emitting regions is important,
and can be compared to observations. The line luminosities, $L$,
shown in the figure correspond to the luminosity in $\LUM$ of photons
escaping the Sobolev region for each line. 
% This is equivalent to $\beta_{ul} n_u A_{ul}$, where the three quantities
% represent the Sobolev escape probability, upper level number density and 
% Einstein A coefficient for the line. 
As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

Figure~\ref{fig:uvspec} shows the synthetic spectrum in the UV from our model. 
To assess the ability of the model to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar
composites from \cite{reichard2003}, normalised to the flux at 2000\AA\
for low inclinations. Unfortunately, the wide variety of
line profile shapes and internal trough structure 
tends to `wash out' BAL troughs to the extent that they do not 
resemble typical BALQSOs.
Because of this, we instead use compare to a {\sl Hubble Space Telescope} 
STIS spectrum of the high BALnicity BALQSO PG0946+301 (Arav et al. 2000),
and an SDSS spectrum of the LoBAL quasar XXXXX (REF),
for the angles of $73^\circ$ and $76^\circ$, respectively. 
We show a cartoon illustrating how geometric effects determine
the output spectra in figure~\ref{fig:sightline}.

@@

\subsection{X-ray Properties and Broadband SEDs}
\label{sec:xray}

@@

\subsection{LoBALs and ionization stratification}

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/ew.png}
\caption
{
$W_{\lambda,RW}$ as a function of inclination in our example model (red line).
We also the mean $1/2$ equivalent width from DiPompeo et al. (2012) as a horizontal
line, with the shaded area corresponding to the standard deviation.
}
\label{fig:ew}
\end{figure}

@@

BALQSOs and quasars generally possess very similar emission line properties 
\citep[e.g.][]{weymann1991,reichard2003}. 
Clearly, the variation with inclination in our models is far greater than 
than the variation between, and standard deviation within, quasar and BALQSO samples.  
This presents a challenge to our model, as well as the geometric unification picture in general.
One obvious potential solution is to hypothesize a more isotropic distribution 
for the emergent condition than predicted by a classical thin disc. 
General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}. 
However, we have verified using AGNSPEC \citep{hubeny2000,hubeny2001,hubenyhubeny1997}
that this effect is small in the UV and optical wavelength regimes, 
and the disc is still anisotropic.

Reprocessing by an extended outflow may also cause a more isotropic
continuum to emerge. Hints that light scattered off a spatially extended 
wind may contribute signiciantly to the emergent continuum come from radiative transfer
simulations \citep{simproga2012} and microlensing observations \citep{sluse2015}.
However, neither of these examples have sufficient reprocessing efficiencies to 
compensate for the disc anisotropy in this case.
An alternative explanation is that the BLR has the same angular distribution 
of emission as the accretion disc.
Indeed, \cite{risaliti2011} find that EW distributions in quasars 
are consistent with anisotropic emission 
from optically thick, disc-like structures for {\em both} the continuum source and BLR. 
If this is the case, it has a dramatic affect on the intrinsic
BAL fraction inferred from flux-limited samples \citep{goodrich1997,krolikvoit1998}.

It is also possible that the equatorial paradigm invoked from early polarisation
studies \citep{goodrich1995, cohen1995,brotherton2006} 
is an over-simplification, or is incorrect. 
High brightness temperatures in some RL BALQSOs imply polar outflows \citep{zhou2006} 
and \cite{bruni2012} find that
RL BALQSOs possess similar radio spectral indices to normal RL quasars, 
suggestive of comparable inclinations. 
In addition, \cite{marin2013} find a bending angle of $\sim45^\circ$ is required to 
explain the polarisation dichotomy of type 1 and 2 AGN using an Elvis-type wind 
model (Elvis 2000). It is therefore possible 
that type 1 quasars and BALQSOs are generally viewed from a fairly narrow range of less
equatorial angles ($\sim0$-$45^\circ$), or that {\em both} 
evolutionary and geometric explanations are required.
We suggest that future modelling should include predictions of polarisation signatures
from a detailed radiative transfer simulation, allowing direct 
comparison with spectropolarimetry of BALQSOs.

>>>>

\title
{
% Modelling quasar outflows:
% clumpy winds and broad emission lines
Testing Quasar Unification: I. Radiative Transfer in Clumpy Winds
%Unifying quasars with clumpy wind models
}

@@

Overall, our work suggests that BAL outflows can produce significant broad emission
lines and that accretion disc winds naturally reproduce the range of
ionization states seen in AGN.
This suggests that geometric unification

@@

pose a specific challenge to models in which an equatorial outflow

@@

In our previous modelling efforts, we have assumed a smooth outflow, 
in which the density at a given point was determined only by the 
kinematic parameters and mass loss rate. However, as already discussed,
AGN winds exhibit significant substructure -- the outflow is expected to be
{\em clumpy}, rather than smooth, and probably on a variety of scales. 
A clumpy outflow offers a possible solution to the so-called `over-ionization problem' in 
quasar and AGN outflows \citep{hamann2013}. This is the main motivation for incorporating 
clumping into our model.

Implementing a treatment of clumping is challenging, for
two main reasons. First, the physical scale lengths and density contrasts 
associated with these parameters are not well-constrained from observations.  
Second, there are significant computational difficulties associated with adequately
resolving and realistically modelling a series of small scale, high density
regions with a MCRT code. 
% Second, the addition of multiple additional degrees of
% freedom in the model results in significantly wider parameter space.
% Unfortunately, the physical scale lengths and density contrasts 
% associated with these parameters are not well-constrained from observations.  
% {\bf KSL: Reverse the order of these difficulties.  The most important is that it is not clear what kind of clumping an AGN wind has; the computational problem is a secondary one JM: Done..}
Given the lack of knowledge about the actual type of clumping, we have adopted
a simple approximation
used successfully in stellar wind modelling, known as 
{\em microclumping} \citep[e.g.][]{hamann1998,hilliermiller1999,hamann2008}. 
The underlying assumption of microclumping is that clump sizes are much smaller than the 
typical photon mean free path, and thus the clumps are 
both geometrically and optically thin. This approach 
allows one to introduce a `volume filling factor', $f_V$. 
The intra-clump medium is assumed to be a vacuum, so the 
density of the clumps is then multiplied by the ``density enhancement'' 
$D=1/f_V$. Opacities, $\kappa$, and emissivities, $\epsilon$, 
can then be expressed as 
\begin{equation}
\kappa = f_V \kappa_C(D);~~\epsilon = f_V \epsilon_C(D).
\end{equation}
Here the subscript $C$ denotes that the quantity is calculated using the 
enhanced density in the clump. The resultant effect is that, {\em for fixed temperature},
 processes that are linear in density, such as electron scattering, are unchanged, 
as $f_V$ and $D$ will cancel out. However, any quantity that scales with the square of density, 
such as collisional excitation or recombination, will increase by a factor of $D$.
In our models, the temperature is not fixed, and is instead set by balancing heating and 
cooling in a given cell. In the presence of an X-ray source, this thermal balance is 
generally dominated by bound-free heating and line cooling. The main effect of including 
clumping in our modelling is that it moderates the ionization state due to the increased 
density. This allows an increase in the ionizing luminosity, amplifying the amount of
bound-free heating and also increasing the competing line cooling term
(thermal line emission).

This treatment is necessarily simple; it does not adequately
represent the complex substructures and stratifications in ionization
state we expect in AGN outflows. Nevertheless, this parameterization allows simple estimates
of the effect clumping might have on the ionization state and emergent 
line emission.

% is that it moderates the ionization state and temperature structure of the plasma

% Clumping the wind has an important effect on the ionization state and has
% been proposed as a solution to the so-called `over-ionization problem' in 
% disc winds \citep{hamann2013}. This is the main motivation for incorporating microclumping
% into our model. This treatment is necessarily simple; it does not adequately
% represent the complex substructures and stratifications in ionization
% state we expect in AGN outflows. Nevertheless, this parameterization allows simple estimates
% of the effect clumping might have on the ionization state and emergent 
% line emission.
% It is also encouraging that microclumping has been used 
% successfully in fits to O-star wind spectra \citep{hillier1991eswingsmodel}.

@@

We then evaluated these models based on whether their synthetic spectra exhibited the 
following properties of quasars and BALQSOs:

\begin{itemize}
\item UV absorption lines 
with $BI > 0$ at $\sim20\%$ of viewing angles (e.g. Knigge et al. 2010)
\item Line emission emerging at low inclinations, with $EW\sim40\AA$ in \civ\ \citep[e.g. ][]{shen2011}
\item H recombination lines with $EW\sim50\AA$ in \la\ \citep[e.g. ][]{shen2011}
\item  \mg\ and \al\ (LoBAL) absorption features with $BI > 0$ at a small subset of BAL angles
\item Verisimilitude with quasar composite spectra.
\end{itemize}
Where $BI$ is the `Balnicity Index' (Weymann et al. 1991), given by
\begin{equation}
BI = \int^{25000}_{3000} \left( 1 - \frac{f(v)}{0.9} \right) dv.
\end{equation}
The constant $C=0$ everywhere, unless the normalized flux
has satisfied $f(v)<0.9$ continuously for at least $2000$~km~s$^{−1}$, 
whereby $C$ is set to $1$.

% the following criteria:
% \begin{itemize}
% \item Do the synthetic spectra exhibit UV absorption lines 
% with $BI > 0$ at $\sim20\%$ of viewing angles (e.g. Knigge et al. 2010)?
% \item Do the synthetic spectra show  line emission emerge at low inclinations, with $EW\sim40\AA$ in \civ \citep[e.g. ][]{shen2011}?
% \item Do H recombination lines appear in the spectra, $EW\sim50\AA$ in \la 
% \citep[e.g. ][]{shen2011}?
% \item Do the spectra produce \mg\ and \al\ (LoBAL) absorption 
% features with $BI > 0$ at a small subset of BAL angles?
% \item Does the model spectra resemble quasar composite spectra?
% \end{itemize}
% Where $BI$ is the `Balnicity Index' (Weymann et al. 1991), given by
% \begin{equation}
% BI = \int^{25000}_{3000} \left( 1 - \frac{f(v)}{0.9} \right) dv.
% \end{equation}
% The constant $C=0$ everywhere, unless the normalized flux
% has satisfied $f(v)<0.9$ continuously for at least $2000$~km~s$^{−1}$, 
% whereby $C$ is set to $1$.

@@

\section{Results From A Fiducial Model}
Here we describe the results from a fiducial model,
and discuss these results in the context of the criteria 
presented in section 3.4. The parameters of this model are shown in table~2.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk. In this section, we examine the physical 
conditions of the flow, and present the synthetic spectra, before comparing
the X-ray properties of this particular model to samples of
quasars and luminous AGN. 
% We also examine trends with inclination in the synthetic spectra, 
% both in terms of the range of ionization states of the absorption lines and equivalent widths of the emission lines.

\begin{table}
\begin{tabular}{p{3cm}p{4cm}}
\hline Fiducial Model Parameters 	&	 Value \\ 
\hline \hline 
$M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
$\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
$\alpha_X$ 	 &	 $-0.9$ \\ 
$L_{X} $ 	 &	 $10^{45}~\rm{erg~s^{-1}}$$^*$ \\ 
$r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
$r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
$\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
$r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
$r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
$\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
$\lambda$ 	&	 $0$ \\ 
$v_{\infty}(r_0)$ 	&	 $v_{esc}(r_0)$ \\ 
$R_v$  	        &	 $10^{19}$cm$^*$ \\ 
$\alpha$ 	&	 $0.6^*$ \\
$f_V$ 	&	 $0.01^*$  \\
$n_x$ 	&	 $100$  \\
$n_z$ 	&	 $200$  \\
\hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
% $\alpha_{OX} $ 	 &	 -2.2\\ 
\end{tabular}
\caption{Wind geometry parameters 
used in the fiducial model, as defined in the text and figure 1.
Parameters differing from the benchmark model of H13 are 
highlighted with an asterisk.}
\label{wind_param}
\end{table}

@@

\noindent
Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clump densities 
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane, 
where $n_H$ is the local number density of H.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity equal to the escape velocity at the streamline base
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration results in
a wind that exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
This is illustrated in figure~\ref{fig:wind} 
by the panels showing the ion fraction $F=n_j/n_{tot}$ of some important ions.
With a clumped wind, we are able to produce the range of ionization states observed
in quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{erg~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

The total line luminosity also increases dramatically compared to H13. 
This is due to heating of the plasma by the increased X-ray 
luminosity, which avoids over-ionization due to clumping.
This line emission typically emerges on the edge of the wind
nearest the central source. The location of the line emitting regions
is dependent on the ionization state, as well as the incident X-rays.
The radii of these emitting regions is important,
and can be compared to observations. The line luminosities, $L$,
shown in the figure correspond to the luminosity in $\LUM$ of photons
escaping the Sobolev region for each line. 
% This is equivalent to $\beta_{ul} n_u A_{ul}$, where the three quantities
% represent the Sobolev escape probability, upper level number density and 
% Einstein A coefficient for the line. 
As shown in figure~\ref{fig:wind},
the \civ\ line in our model is typically formed between 
$100-1000~r_G$ ($\sim10^{17}-10^{18}~\rm{cm}$).
This is in rough agreement with the reverberation mapping 
results of Kaspi (2000) for the $2.6\times10^{9} M_\odot$ quasar S5 0836+71,
and also compares favourably with microlensing measurements of the size of the
\civ\ emission line region in the BALQSO H1413+117 \citep{odowd2015}.

@@

\noindent
Figure~\ref{fig:uvspec} shows the synthetic spectrum in the UV from the fiducial model. 
To assess the ability of the synthetic spectra to match real 
quasar spectra, we also show {\sl Sloan Digital Sky Survey} (SDSS) quasar
composites from \cite{reichard2003}, normalised to the flux at 2000\AA\
for low inclinations. Unfortunately, the wide variety of
line profile shapes and internal trough structure in BALQSOs
tends to `wash out' BAL troughs in composite spectra
to the extent that BALQSO composites do not resemble typical BALQSOs.
Because of this, we instead compare to a {\sl Hubble Space Telescope} 
STIS spectrum of the high BALnicity BALQSO PG0946+301 (Arav et al. 2000),
and an SDSS spectrum of the LoBAL quasar XXXXX (REF),
for the angles of $73^\circ$ and $76^\circ$, respectively. 
We show a cartoon illustrating how geometric effects determine
the output spectra in figure~\ref{fig:sightline}.

@@

\subsection{X-ray Properties}
\label{sec:xray}

@@

\subsection{LoBALs and Ionization Stratification}

@@

\begin{figure}
\centering
\includegraphics[width=0.49\textwidth]{figures/hist_ew.png}
\caption
{
{\sl Top Panel:} A histogram of $1/2$ equivalent widths ($W_{\lambda}/2$)
for BALQSOs and non-BAL quasars from the {\sl SDSS DR7} quasar sample (Shen et al. 2011).
{\sl Bottom Panel:} $W_{\lambda,RW}$ as a function of inclination in our example model (black line).
The green shaded area indicates viewing angles which look into the wind (`BALQSO-like').
}
\label{fig:ew}
\end{figure}

% \begin{figure}
% \centering
% \includegraphics[width=0.45\textwidth]{figures/ew.png}
% \caption
% {
% $W_{\lambda,RW}$ as a function of inclination in our example model (red line).
% We also the mean $1/2$ equivalent width from DiPompeo et al. (2012) as a horizontal
% line, with the shaded area corresponding to the standard deviation.
% }
% \label{fig:ew}
% \end{figure}

@@

BALQSOs and quasars generally possess very similar emission line properties 
\citep[e.g.][]{weymann1991,reichard2003}. This is illustrated in the 
the top panel of figure~\ref{fig:ew}, which shows histograms 
of \civ\ $1/2$ equivalent widths for BAL and non-BAL quasars from the {\sl SDSS DR7} quasar
sample \citep{shen2011}. Clearly the variation in $W_{\lambda,RW}$ at 
BAL-like and quasar-like angles is significantly larger than any systematic 
differences between the quasar and BALQSO samples.
The strong dependence of equivalent width on inclination is caused by decreased continuum
levels at high inclinations. This is due to a combination of
continuum attenuation through the wind, and the effect of limb darkening and foreshortening
in our classical thin disc. This presents a challenge to a geometric unification model
with an equatorial outflow.

% Clearly, the variation with inclination in our models is far greater than 
% than the variation between, and standard deviation within, quasar and BALQSO samples.  
% This presents a challenge to our model, as well as the geometric unification picture in general.
One obvious potential solution is to hypothesize a more isotropic distribution 
for the emergent condition than predicted by a classical thin disc. 
General relativistic effects -- specifically, light bending
and relativistic beaming -- can cause 
the accretion disc SED to become more isotropic \citep[e.g.][]{zhang1997,munozdarias2013}. 
However, we have verified using AGNSPEC \citep{hubeny2000,hubeny2001,hubenyhubeny1997}
that this effect is small in the UV wavelength regime, 
and the disc is still anisotropic.
Reprocessing by an extended outflow may also cause a more isotropic
continuum to emerge. Hints that light scattered off a spatially extended 
wind may contribute signiciantly to the emergent continuum come from radiative transfer
simulations \citep{simproga2012} and microlensing observations \citep{sluse2015}.
However, neither of these examples have sufficient reprocessing efficiencies to 
compensate for the disc anisotropy in this case.

An alternative explanation is that the BLR has the same anisotropic 
angular distribution of emission as the accretion disc.
Indeed, \cite{risaliti2011} find that EW distributions in quasars 
are consistent with anisotropic emission 
from optically thick, disc-like structures for {\em both} the continuum source and BLR. 
If this is the case, it has a dramatic affect on the intrinsic
BAL fraction inferred from flux-limited samples \citep{goodrich1997,krolikvoit1998}.

It is also possible that the equatorial paradigm invoked from early polarisation
studies \citep{goodrich1995, cohen1995,brotherton2006} 
is an over-simplification, or is incorrect. 
High brightness temperatures in some RL BALQSOs imply polar outflows \citep{zhou2006} 
and \cite{bruni2012} find that
RL BALQSOs possess similar radio spectral indices to normal RL quasars, 
suggestive of comparable inclinations. 
In addition, \cite{marin2013} find a bending angle of $\sim45^\circ$ is required to 
explain the polarisation dichotomy of type 1 and 2 AGN using an Elvis-type wind 
model (Elvis 2000). It is therefore possible 
that type 1 quasars and BALQSOs are generally viewed from a fairly narrow range of less
equatorial angles ($\sim0$-$45^\circ$), or {\em both} 
evolutionary and geometric explanations are required.
We suggest that future modelling should include predictions of polarisation signatures
from a detailed radiative transfer simulation, allowing direct 
comparison with spectropolarimetry of BALQSOs.
